---
title: Functional niches of cleanerfish species are mediated by habitat use, cleaning intensity, and client selectivity
author: "Simon J Brandl"
date: "09/05/2020"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---

# Functional niches of cleanerfish species are mediated by habitat use, cleaning intensity, and client selectivity

This repository contains the code and data to reproduce all tables and
figures presented in Côté & Brandl “Functional niches of cleanerfish species are mediated by habitat use, cleaning intensity, and client selectivity.” The abstract is pasted below: 

1.	An animal’s functional niche is a complex, multidimensional construct, mediated by an individual’s morphology, physiology, and behaviour. Behavioural aspects of the niche can be difficult to quantify, as their expression is often subtle and tailored to an infinite number of different situations that involve sophisticated mechanisms such as mutualisms, species dominance, or fear effects. 
2.	The extreme diversity of tropical fish assemblages has led to extensive debate over the extent to which species differ in their resource use and functional role. Ectoparasite removal by cleanerfish species is considered a behaviourally complex interspecific interactions in vertebrates, but differences in the services rendered by various species of cleanerfish, and potential consequences for the range of clients (i.e., resources) they attract, have rarely been examined. 
3.	Here, we quantify differences among three coexisting species of morphologically similar cleaner wrasses (Labroides bicolor, L. dimidiatus, and L. pectoralis) in the global centre of marine biodiversity, the Coral Triangle. 
4.	We found no clear taxonomic partitioning of clients among cleanerfishes. However, the three cleanerfish species exhibited distinct habitat preferences, and differed in their cleaning intensity: L. bicolor serviced the fewest species and clients, while L. pectoralis serviced the most clients and spent the most time cleaning. 
5.	Accordingly, L. pectoralis showed no preference for clients based on client size or abundance, while both L. bicolor and L. dimidiatus had a higher likelihood of interacting with clients based on their size (larger client species in L. bicolor, smaller client species in L. dimidiatus) and abundance (more abundant client species for both).
6.	Our results suggest that the services rendered by the three species of cleanerfishes differ in their spatial availability, quality, and selectivity, thus permitting the coexistence of these species despite their ecological similarity. This, in turn, creates a complex seascape of species-specific cleaning services that underpins crucial biotic interactions in the ocean’s most diverse ecosystem.


## Instructions
All analyses for the current project were done in R, using the drake pipeline [drake](https://github.com/ropensci/drake). You can use drake to compile all models, figures, and tables.
To do so, first install `drake` from CRAN: 

```{r, eval=FALSE}
install.packages("drake")
```

Next you need to open an R session with working directory set to the root of the project.    

We use a number of packages, listed in `r/packages.R`. If needed, missing packages should first be installed. 

Then, to generate all figures, analyses, and tables, simply run:

```{r, eval = FALSE}
drake::r_make()
```

All output will be automatically rendered inside the folder called output. 

## Details

The elements of this project include:

1.  data: this folder includes all the raw data necessary to run the
    script.
2.  output: this folder includes all outputs from the R-script.
3.  r: this folder contains three .R files, `packages.R`, `functions.R`,
    and `plan.R`.  
    `packages.R` contains all packages needed, `functions.R` contains
    all functions to reproduce the analyses, and `plan.R` provides the
    code that binds each step of the workflow.

## Datasets

1.  depth (labroides_depth_sjb.csv): Contains depth distributions of three cleanerfish species on two reefs in Indonesia. Each row is an observation, columns contain data on site (Pinnacle or Hoga Home Reef), depth, and species.

2.  habitat (labroides_habitat_sjb.csv): Contains information about habitat properties of territories and home ranges of the three Labroides species. Rows are specific cleaner wrasse individuals. Columns 1-4 contain the station (cleaner wrasse individuals), species (1 = L. dimidiatus, 2 = L. pectoralis, 3 = L. bicolor), life stage (age = juvenile or adult), and the number of fish present at a given station (no_fish; only one individual observed per station). Columns 5-9 provide detailed benthic compositional data (relative cover), while columns 10 and 11 provide summarized proportions of live vs. dead habitat. Column 12 provides territory size in square meters and column 13 the depth of the specific territory.

3.  temp (labroides_temporal_sjb.csv): Contains data on temporal patterns in cleaning interactions. Rows are cleaning stations (replicates). Columns 1-3 provide metadata as previously, while the remaining columns provide the three interaction variables across the three time periods. spp = Number of client species; client = Number of clients; time = Seconds spent inspecting/cleaning; am = 7h00-9h00; noon = 11h00-13h00; pm = 15h00-17h00. 

4.  Cleaning interaction datasets:
    a.  clean.interac (labroides_cleaninginteractions_sjb.csv):  
        full dataset of interactions between client species (common name in column 1) and cleaner fishes (columns). Column names are station names, which can be cross-matched with species in the "stations" dataset. 
    b.  clean.clients (labroides_clients_meta_sjb.csv): 
        metadata containing scientific and common names of all clients, as well as their families.
    c.  clean.stations (.coi = COI primer data)labroides_sites_meta_sjb.csv):
        metadata containing species affiliation for each station ID. 1 = L. dimidiatus, 2 = L. pectoralis, 3 = L. bicolor.
    d.  fishID.all (fishID.merged.all.csv): 
        metadata containing common and scientific names of all mobile fish species present at Hoga Home reef. 
    e.  fish.abun (labroides_fishcounts_sjb.csv):
        abundances of all fish species in the area, as determined by point counts. First column gives species names, rest          of the columns are one point count each.
    f.  client.size (labroides_client_sizes.csv):
        maximum size of all client species as obtained from FishBase.

## Supplemental info

This paper was produced using the following software and associated packages:

```{r, echo=FALSE}
devtools::session_info()
```

All code written by Simon J. Brandl
(<a href="mailto:simonjbrandl@gmail.com" class="email">simonjbrandl@gmail.com</a>
and
<a href="https://github.com/simonjbrandl" class="uri">https://github.com/simonjbrandl</a>).
Please contact me for any issue or question.
